Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles

### 2022-12-22 Manuscript_rev2-LL.tex
more minor edits
more minor updates

\begin{abstract}
  % \justifying
  %
  While optical trapping is a well understood method for force
  transduction and detection,characterisation of trapped entities
  poses a two-fold challenge --- one experimental concerning the
  arrangement of light detectors and the other, theoretical involving
  solving of the inverse light scattering problem. Combining static
  light scattering techniques with optical trapping poses significant
  engineering challenges due to the space constraints in a
  conventional optical trapping setup.  We propose here a plausible
  scenario of detecting scattered light from an optically trapped
  asymmetric microstructure using a novel, multi-angle, optical-fibre
  based detection scheme and demonstrate how a Bayesian inference
  based analysis of the data simulated to mimic light scattering
  detection signals in such scenarios maybe used for solving the
  inverse light scattering problem and help characterise trapped
  entities.  To this end, we discuss the application of our method to
  infer the instantaneous orientations of an asymmetric microsphere
  dimer being trapped. We argue that this method can be extended for
  determining any characteristics of the trapped microstructure that
  influence the light scattering pattern.
\end{abstract}

@@

To measure the error in the model's estimate, the distribution from
Eq.~\eqref{eq:bayes} and an ideal result, $p_{ideal}$, were compared.
An ideal result would be one that is 0 for every $\hat{\bf n}$ apart
from $\hat{\bf n}_{best}$.
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation. The model's error was
quantified by calculating the Kullback-Leibler divergence $K_l$
between the two distributions:
\begin{align}
K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
\right]
\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less
confident in its prediction of the dimer's orientation. While the
divergence calculations are useful for showing our model's confidence
in the prediction it cannot accurately reflect if the model is correct
or not with its estimation.  Section~\ref{sec:Bayes} and
Section~\ref{sec:divergence} are all for a single measurement of a
light scattering signal from our dimer.  To properly model it's motion
we need to consider how the orientation changes with time.

@@

The worst case scenario is akin to randomly choosing a reference orientation at each time step. The greater the value of $F(K_l)$ the better our model's confidence is in characterising the dimer's motion. Because our model is dependent on several parameters we need to a sophisticated method for understanding how these parameters correlate with $F(K_l)$.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width =\textwidth]{./Images/fig5a.png}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig5b.png}
\end{subfigure}
\caption{\label{fig:motion}
%
Simulation results of: (a) the dimers orientation vector with time,
(b) the dimer's [x,y,z] position with time.
%
}
\end{figure}

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.  Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.  It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely. Further simulations of differently sized dimers showed a similar behaviour of moving away from the trap focus while remaining trapped, but only when $a_1 \geq 2a_2$.  Dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
%
In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations. Once we have our experimental set up complete we can confirm this by trapping an asymmetric dimer and monitoring its orientation.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.32\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig6a.png}
\end{subfigure}
\begin{subfigure}{0.32\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig6b.png}
\end{subfigure}
\begin{subfigure}{0.32\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig6c.png}
\end{subfigure}
\caption{\label{fig:epsilon}
%
Model prediction for signal error of (a) $1\%$ ~ $[F(K_l)=1.174]$, (b)
$15\%$ ~ [$[F(K_l)=0.606]$, and (c) $25\%$ ~ $[F(K_l)=0.457]$.
%
}
\end{figure}

As can be seen from Figure~\ref{fig:epsilon}, the inclusion of a signal noise quickly leads to a decrease in our model's performance. This is due to an inherent feature of the inverse scattering problem; that being that a two distinct regions of orientation space will become heavily intertwined in the intensity space despite remaining continuous. As the signal error increases our neural network will map the signal to the incorrect reference orientation.
%
Fortunately, we can alleviate this by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap. We repeated the model at for the above signal errors, but increasing $\beta$ until the model best fit our dimer's motion

>>>>

\begin{abstract}
  % \justifying
  %
  While optical trapping is a well understood method for force
  transduction and detection, characterisation of trapped entities
  poses a two-fold challenge --- one experimental, concerning the
  arrangement of light detectors and the other, theoretical involving
  solving of the inverse light scattering problem. Combining static
  light scattering techniques with optical trapping poses significant
  engineering challenges due to the space constraints in a
  conventional optical trapping setup.  We propose a plausible
  scenario of detecting scattered light from an optically trapped
  asymmetric microstructure using a novel, multi-angle, optical-fibre
  based detection scheme and demonstrate how a Bayesian inference
  based analysis of the data simulated to mimic light scattering
  detection signals in such scenarios maybe used for solving the
  inverse light scattering problem and help characterise trapped
  entities.  To this end, we discuss the application of our method to
  infer the instantaneous orientations of an asymmetric microsphere
  dimer being trapped. We argue that this method can be extended for
  determining any characteristics of the trapped microstructure that
  influence the light scattering pattern.
\end{abstract}

@@

To measure the error in the model's estimate, the distribution from
Eq.~\eqref{eq:bayes} and an ideal result, $p_{ideal}$, were compared.
An ideal result would be one that is 0 for every $\hat{\bf n}$ apart
from $\hat{\bf n}_{best}$.
\begin{align}
\label{eq:best}
p_{best} = 
\begin{cases}
1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$} \\
0 & \text{anywhere else}
\end{cases}
\end{align}
The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation. The model's error was
quantified by calculating the Kullback-Leibler divergence $K_l$
between the two distributions:
\begin{align}
K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
\right]
\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less
confident in its prediction of the dimer's orientation. While the
divergence calculations are useful for showing our model's confidence
in the prediction it cannot accurately reflect if the model is correct
or not with its estimation.  Section~\ref{sec:Bayes} and
Section~\ref{sec:divergence} are all for a single measurement of a
light scattering signal from our dimer.  To properly model it's motion
we need to consider how the orientation changes with time.

@@

The worst case scenario is akin to randomly choosing a reference
orientation at each time step. The greater the value of $F(K_l)$, the
better our model's confidence is in characterising the dimer's
motion. Because our model is dependent on several parameters we need
to a sophisticated method for understanding how these parameters
correlate with $F(K_l)$.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width =\textwidth]{./Images/fig5a.png}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig5b.png}
\end{subfigure}
\caption{\label{fig:motion}
%
Simulation results for the dimer (a) orientation and
(b) position with time.
%
}
\end{figure}

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.  Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.  It is interesting to note that the dimer is furthest from the trap centre as it goes into a horizontal orientation before drawing closer again as it inverts completely. Further simulations of differently sized dimers showed a similar behaviour of moving away from the trap focus while remaining trapped, but only when $a_1 \geq 2a_2$.  Dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
%
In the simulations of Vigilante~\emph{et~al}.\
\cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were
investigated; their findings showed that the optical torque on the
dimer goes to zero while aligned vertically and is at its maximum in a
horizontal alignment. Therefore, the inversion of an asymmetric dimer
suggests that if the size difference is significant, the optical
torque is minimal for a dimer in both horizontal and vertical
orientations. Once we have our experimental set up complete we can
confirm this by trapping an asymmetric dimer and monitoring its
orientation.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.32\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig6a.png}
\end{subfigure}
\begin{subfigure}{0.32\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig6b.png}
\end{subfigure}
\begin{subfigure}{0.32\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth]{./Images/fig6c.png}
\end{subfigure}
\caption{\label{fig:epsilon}
%
Model prediction for signal error of (a) $\eta=1\%$ ~
$[F(K_l)=1.174]$, (b) $\eta=15\%$ ~ [$[F(K_l)=0.606]$, and (c)
$\eta=25\%$ ~ $[F(K_l)=0.457]$.
%
}
\end{figure}

As can be seen from Figure~\ref{fig:epsilon}, the inclusion of noise
quickly leads to a decrease in the method's performance.  This is due
to an inherent feature of the inverse scattering problem: two distinct
regions of orientation space will become highly intertwined in the
intensity space despite each remaining continuously connected.  As the
signal error increases our neural network will map the signal to the
incorrect reference orientation.
%
Fortunately, we can alleviate this by increasing the weighting factor $\beta$ that reflects the degree of freedom our dimer has within the trap. We repeated the model at for the above signal errors, but increasing $\beta$ until the model best fit our dimer's motion
